---
title: "Hands_on Exercise 2: Global and Local Measures of Spatial Autocorrelation"
editor: visual
---

## Overview

In this hands-on exercise, I learn how to compute Global and Local Measure of Spatial Autocorrelation (GLSA) by using **spdep** package.. The data sets are extracted from the following sources:

-   Hunan province administrative boundary layer at county level. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan's local development indicators in 2012.

## Getting Started

The code chunk below install and load **spdep**, **sf**, **tmap** and **tidyverse** packages of R. **sf** package would be used for importing, managing, and processing geospatial data, and **tidyverse** package would be used for performing data science tasks such as importing, wrangling and visualising data. **pdep** will be used to compute spatial weights, global and local spatial autocorrelation statistics, and **tmap** will be used to prepare cartographic quality chropleth map.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## Getting the Data Into R Environment

### Import shapefile into r environment

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

The code below imports *Hunan_2012.csv* into R by using *read_csv()* of **readr** package. The output is R data frame class.

```{r, message=FALSE}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### Performing relational join

The code chunk below will be used to update the attribute table of *hunan*'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. This is performed by using *left_join()* of **dplyr** package.

```{r, message=FALSE}
hunan <- left_join(hunan,hunan2012)
```

### Visualising Regional Development Indicator

The code below prepares a basemap and a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r}
equal <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal interval classification")

quantile <- tm_shape(hunan) +
  tm_fill("GDPPC",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(main.title = "Equal quantile classification")

tmap_arrange(equal, 
             quantile, 
             asp=1, 
             ncol=2)
```

## Global Spatial Autocorrelation

### Computing Contiguity Spatial Weights

Before computing the global spatial autocorrelation statistics, The spatial weights are required and used to define the neighbourhood relationships between the geographical units (i.e. county) in the study area.

[*poly2nb()*](https://r-spatial.github.io/spdep/reference/poly2nb.html) of **spdep** package is used to compute contiguity weight matrices for the study area. More specifically, the code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, 
                queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are 88 area units in Hunan. The most connected area unit has 11 neighbours. There are two area units with only one heighbours.

### Row-standardised weights matrix

The code chunk below assigns equal weights to each neighboring polygon (style="W"). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. Other more robust options are available, notably style="B".

```{r}
rswm_q <- nb2listw(wm_q, 
                   style="W", 
                   zero.policy = TRUE)
rswm_q
```

The input of *nb2listw()* must be an object of class **nb**. The syntax of the function has two major arguments, namely style and zero.poly.

-   *style* can take values "W", "B", "C", "U", "minmax" and "S". B is the basic binary coding, W is row standardised (sums over all links to n), C is globally standardised (sums over all links to n), U is equal to C divided by the number of neighbours (sums over all links to unity), while S is the variance-stabilizing coding scheme proposed by Tiefelsdorf.

-   If *zero policy* is set to TRUE, weights vectors of zero length are inserted for regions without neighbour in the neighbours list. These will in turn generate lag values of zero, equivalent to the sum of products of the zero row t(rep(0, length=length(neighbours))) %\*% x, for arbitrary numerical vector x of length length(neighbours). The spatially lagged value of x for the zero-neighbour region will then be zero, which may (or may not) be a sensible choice.

### Global Spatial Autocorrelation: Moran's I

### Maron's I test

The code chunk below performs Moran's I statistical testing using [*moran.test()*](https://r-spatial.github.io/spdep/reference/moran.test.html) of **spdep**.

```{r}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```

> From the output above, we can conclude that null hypothesis (regions are spatially independent) should be rejected as p-value is smaller than 0.05. As z-value is larger than 0, it indicates that the regions are clustered, observations tend to be similar.

#### Computing Monte Carlo Moran's I

The code chunk below performs permutation test for Moran's I statistic by using [*moran.mc()*](https://r-spatial.github.io/spdep/reference/moran.mc.html) of **spdep**. A total of 1000 simulation will be performed.

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, 
                listw=rswm_q, 
                nsim=999, 
                zero.policy = TRUE, 
                na.action=na.omit)
bperm
```

> From the output above, we can conclude that null hypothesis (regions are spatially independent) should be rejected as p-value is smaller than 0.05. As z-value is larger than 0, it indicates that the regions are clustered, observations tend to be similar.

#### 

#### Visualising Monte Carlo Moran's I

The code chunk below examines the simulated Moran's I test statistics in greater detail by plotting the distribution of the statistical values as a histogram using hist() and [*abline()*](https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/abline) of R Graphics.

```{r}
mean(bperm$res[1:999])
var(bperm$res[1:999])
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, 
     freq=TRUE, 
     breaks=20, 
     xlab="Simulated Moran's I")
abline(v=0, 
       col="red") 
```

> From the output above, it could be seen that the distribution of residuals or error terms is positively skewed, indicating that there is spatial autocorrelation among the regions.

### Global Spatial Autocorrelation: Geary's

#### Geary's C test

The code chunk below performs Geary's C test for spatial autocorrelation by using [*geary.test()*](https://r-spatial.github.io/spdep/reference/geary.test.html) of **spdep**.

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```

> From the output above, we can conclude that there is no significant spatial autocorrelation as p-value is larger than 0.05 (fail to reject null hypothesis).

#### Computing Monte Carlo Geary's C

The code chunk below performs permutation test for Geary's C statistic by using [*geary.mc()*](https://r-spatial.github.io/spdep/reference/geary.mc.html) of **spdep**.

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, 
               listw=rswm_q, 
               nsim=999)
bperm
```

> From the output above, we can conclude that there is no significant spatial autocorrelation as p-value is larger than 0.05 (fail to reject null hypothesis).

#### Visualising the Monte Carlo Geary's C

The code chunk below plost a histogram to reveal the distribution of the simulated values.

```{r}
mean(bperm$res[1:999])
var(bperm$res[1:999])
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

> From the output, we can see that residuals or error terms are independent and normally distributed, thereby validating the null hypothesis that values at one location do not depend on values at other (neighbouring) locations.

## Spatial Correlogram

Spatial correlograms are great to examine patterns of spatial autocorrelation in your data or model residuals. They show how correlated are pairs of spatial observations when you increase the distance (lag) between them - they are plots of some index of autocorrelation (Moran's I or Geary's c) against distance.

### Compute Moran's I correlogram

In the code chunk below, [*sp.correlogram()*](https://r-spatial.github.io/spdep/reference/sp.correlogram.html) of **spdep** package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Moran's I. The **plot()** of base graph is then used to plot the output.

```{r}
MI_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="I", 
                          style="W")
plot(MI_corr)
```

The code chunk below allows us to examine the full analysis report and have a complete interpretation.

```{r}
print(MI_corr)
```

> From the report above, we can conclude that the null hypothesis that pairs of spatial observations are correlated is rejected as the p-value is smaller than 0.05.

### Compute Geary's C correlogram and plot

In the code chunk below, *sp.correlogram()* of **spdep** package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Geary's C. The **plot()** of base Graph is then used to plot the output.

```{r}
GC_corr <- sp.correlogram(wm_q, 
                          hunan$GDPPC, 
                          order=6, 
                          method="C", 
                          style="W")
plot(GC_corr)
```

Similar to the previous step, we will print out the analysis report by using the code chunk below.

```{r}
print(GC_corr)
```

## Cluster and Outlier Analysis

Local Indicators of Spatial Association or LISA are statistics that evaluate the existence of clusters in the spatial arrangement of a given variable.

### Computing local Moran's I

To compute local Moran's I, the [*localmoran()*](https://r-spatial.github.io/spdep/reference/localmoran.html) function of **spdep** will be used. It computes *Ii* values, given a set of *zi* values and a listw object providing neighbour weighting information for the polygon associated with the zi values.

The code chunks below are used to compute local Moran's I of *GDPPC2012* at the county level.

```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```

*localmoran()* function returns a matrix of values whose columns are:

-   Ii: the local Moran's I statistics

-   E.Ii: the expectation of local moran statistic under the randomisation hypothesis

-   Var.Ii: the variance of local moran statistic under the randomisation hypothesis

-   Z.Ii:the standard deviate of local moran statistic

-   Pr(): the p-value of local moran statistic

The code chunk below list the content of the local Moran matrix derived by using [*printCoefmat()*](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/printCoefmat).

```{r}
printCoefmat(data.frame(localMI[fips,], row.names=hunan$County[fips]), check.names=FALSE)
```

#### Mapping the local Moran's I

The code chunks below appends the local Moran's I dataframe (i.e. localMI) onto hunan SpatialPolygonDataFrame. The out SpatialPolygonDataFrame is called *hunan.localMI*.

```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

#### Mapping local Moran's I values

Using choropleth mapping functions of **tmap** package, we can plot the local Moran's I values by using the code chinks below.

```{r, message=FALSE}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)
```

#### Mapping local Moran's I p-values

The code chunks below produce a choropleth map of Moran's I p-values by using functions of **tmap** package.

```{r}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

```

#### Mapping both local Moran's I values and p-values

The code chunk below plot both the local Moran's I values map and its corresponding p-values map next to each other for effective interpretation.

```{r, message=FALSE}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

## Creating a LISA Cluster Map

The LISA Cluster Map shows the significant locations color coded by type of spatial autocorrelation. The first step before we can generate the LISA cluster map is to plot the Moran scatterplot.

### Plotting Moran scatterplot

The Moran scatterplot is an illustration of the relationship between the values of the chosen attribute at each location and the average value of the same attribute at neighboring locations.

The code chunk below plots the Moran scatterplot of GDPPC 2012 by using [*moran.plot()*](https://r-spatial.github.io/spdep/reference/moran.plot.html) of **spdep**.

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```

Notice that the plot is split in 4 quadrants. The top right corner belongs to areas that have high GDPPC and are surrounded by other areas that have the average level of GDPPC. This are the high-high locations in the lesson slide.

### Plotting Moran scatterplot with standardised variable

The code chunk below uses [*scale()*](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/scale) to centers and scales the variable.

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% as.vector 
```

The [*as.vector()*](https://www.rdocumentation.org/packages/pbdDMAT/versions/0.5-1/topics/as.vector) added to the end is to make sure that the resultant data type is a vector, that map neatly into out dataframe.

The code chunk below plots the Moran scatterplot.

```{r}
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

### Preparing LISA map classes

The code chunks below prepares a LISA cluster map.

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
DV <- hunan$GDPPC - mean(hunan$GDPPC)     
C_mI <- localMI[,1] - mean(localMI[,1])    
signif <- 0.05       
quadrant[DV >0 & C_mI>0] <- 4      
quadrant[DV <0 & C_mI<0] <- 1      
quadrant[DV <0 & C_mI>0] <- 2
quadrant[DV >0 & C_mI<0] <- 3
quadrant[localMI[,5]>signif] <- 0
```

### Plotting LISA map

The code chunk below builds the LISA map.

```{r}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```

The code chunk below plot both the local Moran's I values map and its corresponding p-values map next to each other for effective interpretation.

```{r, message=FALSE}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, asp=1, ncol=2)
```

> From the LISA map above, we can conclude that regions with high GDPPC generally are highly correlated with their neighbors, hence forming clusters that are similar.

## Hot Spot and Cold Spot Area Analysis

Beside detecting cluster and outliers, localised spatial statistics can be also used to detect hot spot and/or cold spot areas. The term 'hot spot' has been used generically across disciplines to describe a region or value that is higher relative to its surroundings.

### Getis and Ord's G-Statistics

An alternative spatial statistics to detect spatial anomalies is the Getis and Ord's G-statistics. It looks at neighbours within a defined proximity to identify where either high or low values clutser spatially.

The analysis consists of three steps:

-   Deriving spatial weight matrix

-   Computing Gi statistics

-   Mapping Gi statistics

### Deriving distance-based weight matrix

For Getis-Ord, neighbours are defined based on distance and there are two type of distance-based proximity matrix: fixed distance weight matrix; and adaptive distance weight matrix.

#### Deriving the centroid

The code chunk below uses map_dbl variation of map from the purrr package to map the *st_centroid()* function over the geometry column of us.bound and access the longitude and latitiude values in each centroid.

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
```

#### Determine the cut-off distance

The code chunk below determines the upper limit for distance band by using the steps below:

-   Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using [*knearneigh()*](https://r-spatial.github.io/spdep/reference/knearneigh.html) of **spdep**.

-   Convert the knn object returned by *knearneigh()* into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using [*knn2nb()*](https://r-spatial.github.io/spdep/reference/knn2nb.html).

-   Return the length of neighbour relationship edges by using [*nbdists()*](https://r-spatial.github.io/spdep/reference/nbdists.html) of **spdep**. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.

-   Remove the list structure of the returned object by using [**unlist()**](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/unlist).

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

#### Computing fixed distance weight matrix

The code chunk below computes the distance weight matrix by using [*dnearneigh()*](https://r-spatial.github.io/spdep/reference/dnearneigh.html).

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

The code chunk belows uses *nb2listw()* to convert the nb object into spatial weights object.

```{r}
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

The output spatial weights object is called `wm62_lw`.

### Computing adaptive distance weight matrix

One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours.

The code chunk below controls the numbers of neighbours directly using k-nearest neighbours.

```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn
```

Next, *nb2listw()* is used to convert the nb object into spatial weights object.

```{r}
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)
```

## Computing Gi statistics

### Gi statistics using fixed distance

```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
gi.fixed
```

The output of localG() is a vector of G or Gstar values, with attributes "gstari" set to TRUE or FALSE, "call" set to the function call, and class "localG". The Gi statistics is represented as a Z-score. Greater values represent a greater intensity of clustering and the direction (positive or negative) indicates high or low clusters.

The code chunk below joins the Gi values to their corresponding hunan sf data frame. It convert the output vector (i.e. *gi.fixed*) into r matrix object by using *as.matrix()*. Next, *cbind()* is used to join hunan\@data and *gi.fixed* matrix to produce a new SpatialPolygonDataFrame called *hunan.gi*. Lastly, the field name of the gi values is renamed to *gstat_fixed* by using *rename()*.

```{r}
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```

### Mapping Gi values with fixed distance weights

The code chunk below shows the functions used to map the Gi values derived using fixed distance weight matrix.

```{r, message=FALSE}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

> From the Gi map above, we can conclude that regions with high GDPPC are generally hot spot areas with significant and positive Gi (the region is associated with relatively high values in surrounding locations). Regions with low GDPPC are generally cold spot areas with significant and negative Gi (the region is associated with relatively low values in surrounding locations).

### Gi statistics using adaptive distance

The code chunk below are used to compute the Gi values for GDPPC2012 by using an adaptive distance weight matrix (i.e *knb_lw*).

```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
```

### Mapping Gi values with adaptive distance weights

The code chunk below uses the choropleth mapping functions of **tmap** package to map the Gi values derived using fixed distance weight matrix.

```{r, message=FALSE}
gdppc<- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) + 
  tm_fill(col = "gstat_adaptive", 
          style = "pretty", 
          palette="-RdBu", 
          title = "local Gi") + 
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, 
             Gimap, 
             asp=1, 
             ncol=2)
```

> From the Gi map above, we can conclude that regions that are nearer to each other are generally hot spot areas with significant and positive Gi (the region is associated with relatively high values in surrounding locations). Regions that are further apart are generally cold spot areas with significant and negative Gi (the region is associated with relatively low values in surrounding locations).
