---
title: "Hands_on_Ex4"
editor: visual
execute: 
  warning: false
  message: false
  fig-retine: 3
---

## Overview

**Geographically weighted regression (GWR)** is a spatial statistical technique that takes non-stationary variables into consideration (e.g., climate; demographic factors; physical environment characteristics) and models the local relationships between these independent variables and an outcome of interest (also known as dependent variable).

In this hands-on exercise, I will practice how to build [hedonic pricing](https://www.investopedia.com/terms/h/hedonicpricing.asp) models by using GWR methods. The dependent variable is the resale prices of condominium in 2015. The independent variables are divided into either structural and locational.

## The Data

Two data sets will be used in this model building exercise, they are:

-   URA Master Plan subzone boundary in shapefile format (i.e. *MP14_SUBZONE_WEB_PL*)

-   condo_resale_2015 in csv format (i.e. *condo_resale_2015.csv*)

## Getting Started

The R packages needed for this exercise are as follows:

-   R package for building OLS and performing diagnostics tests

    -   [**olsrr**](https://olsrr.rsquaredacademy.com/)

-   R package for calibrating geographical weighted family of models

    -   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/)

-   R package for multivariate data visualisation and analysis

    -   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html)

-   Spatial data handling

    -   **sf**

-   Attribute data handling

    -   **tidyverse**, especially **readr**, **ggplot2** and **dplyr**

-   Choropleth mapping

    -   **tmap**

The code chunks below installs and launches these R packages into R environment.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

[**GWmodel**](https://www.jstatsoft.org/article/view/v063i17) package provides a collection of localised spatial statistical methods, namely: GW summary statistics, GW principal components analysis, GW discriminant analysis and various forms of GW regression; some of which are provided in basic and robust (outlier resistant) forms. Commonly, outputs or parameters of the GWmodel are mapped to provide a useful exploratory tool, which can often precede (and direct) a more traditional or sophisticated statistical analysis.

## Geospatial Data Wrangling

### Importing geospatial data

The geospatial data used in this hands-on exercise is called MP14_SUBZONE_WEB_PL. It is in ESRI shapefile format. The shapefile consists of URA Master Plan 2014's planning subzone boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import *MP_SUBZONE_WEB_PL* shapefile by using `st_read()` of **sf** packages.

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

The report above shows that the R object used to contain the imported MP14_SUBZONE_WEB_PL shapefile is called *mpsz* and it is a simple feature object. The geometry type is *multipolygon*. Note that mpsz simple feature object does not have EPSG information.

### Updating CRS information

The code chunk below updates the newly imported *mpsz* with the correct ESPG code (i.e. 3414)

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

The code chunk below verifies the projection of the newly transformed *mpsz_svy21* by using `st_crs()` of **sf** package.

```{r}
st_crs(mpsz_svy21)
```

Notice that the EPSG is indicated as *3414* now.

Next, the code chunk below will reveal the extent of *mpsz_svy21* by using `st_bbox()` of sf package.

```{r}
st_bbox(mpsz_svy21) 
```

## Aspatial Data Wrangling

### Importing the aspatial data

The *condo_resale_2015* is in csv file format. The codes chunk below uses `read_csv()` function of **readr** package to import *condo_resale_2015* into R as a tibble data frame called *condo_resale*.

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

The codes chunks below uses `glimpse()` to display the data structure of the file and examine if the data file has been imported correctly.

```{r}
glimpse(condo_resale)
```

```{r}
head(condo_resale$LONGITUDE) #see the data in XCOORD column
head(condo_resale$LATITUDE) #see the data in YCOORD column
```

Next, `summary()` of base R is used to display the summary statistics of *cond_resale* tibble data frame.

```{r}
summary(condo_resale)
```

### Converting aspatial data frame into a sf object

Currently, the *condo_resale* tibble data frame is aspatial. The code chunk below converts condo_resale data frame into a simple feature data frame by using `st_as_sf()` of **sf** packages.

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

`st_transform()` of **sf** package is used to convert the coordinates from wgs84 (i.e. crs:4326) to svy21 (i.e. crs=3414). Next, `head()` is used to list the content of *condo_resale.sf* object.

```{r}
head(condo_resale.sf)
```

Notice that the output is in point feature data frame.

## Exploratory Data Analysis (EDA)

### EDA using statistical graphics

The code chunk below plots the distribution of *SELLING_PRICE* by using appropriate Exploratory Data Analysis (EDA).

```{r}
ggplot(data=condo_resale.sf, aes(x=`SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

The figure above reveals a right skewed distribution. This means that more condominium units were transacted at relative lower prices.

Statistically, the skewed dsitribution can be normalised by using log transformation. The code chunk below is used to derive a new variable called *LOG_SELLING_PRICE* by using a log transformation on the variable *SELLING_PRICE*. It is performed using `mutate()` of **dplyr** package.

```{r}
condo_resale.sf <- condo_resale.sf %>%
  mutate(`LOG_SELLING_PRICE` = log(SELLING_PRICE))
```

The code chunk below plots the *LOG_SELLING_PRICE*.

```{r}
ggplot(data=condo_resale.sf, aes(x=`LOG_SELLING_PRICE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

Notice that the distribution is relatively less skewed after the transformation.

### Multiple Histogram Plots distribution of variables

The code chunk below is used to create 12 histograms. Then, `ggarrange()` is used to organised these histogram into a 3 columns by 4 rows small multiple plot.

```{r, fig.width=12, fig.height=8}
AREA_SQM <- ggplot(data=condo_resale.sf, aes(x= `AREA_SQM`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

AGE <- ggplot(data=condo_resale.sf, aes(x= `AGE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CBD <- ggplot(data=condo_resale.sf, aes(x= `PROX_CBD`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_CHILDCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_CHILDCARE`)) + 
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_ELDERLYCARE <- ggplot(data=condo_resale.sf, aes(x= `PROX_ELDERLYCARE`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_URA_GROWTH_AREA <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_URA_GROWTH_AREA`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_HAWKER_MARKET <- ggplot(data=condo_resale.sf, aes(x= `PROX_HAWKER_MARKET`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_KINDERGARTEN <- ggplot(data=condo_resale.sf, aes(x= `PROX_KINDERGARTEN`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_MRT <- ggplot(data=condo_resale.sf, aes(x= `PROX_MRT`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PARK <- ggplot(data=condo_resale.sf, aes(x= `PROX_PARK`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_PRIMARY_SCH <- ggplot(data=condo_resale.sf, aes(x= `PROX_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

PROX_TOP_PRIMARY_SCH <- ggplot(data=condo_resale.sf, 
                               aes(x= `PROX_TOP_PRIMARY_SCH`)) +
  geom_histogram(bins=20, color="black", fill="light blue")

ggarrange(AREA_SQM, AGE, PROX_CBD, PROX_CHILDCARE, PROX_ELDERLYCARE, 
          PROX_URA_GROWTH_AREA, PROX_HAWKER_MARKET, PROX_KINDERGARTEN, PROX_MRT,
          PROX_PARK, PROX_PRIMARY_SCH, PROX_TOP_PRIMARY_SCH,  
          ncol = 3, nrow = 4)
```

### Drawing Statistical Point Map

Lastly, I would reveal the geospatial distribution condominium resale prices in Singapore by using **tmap** package.

The code chunk below turns on the interactive mode of tmap.

```{r}
tmap_mode("view")
sf::sf_use_s2(FALSE)
```

Next, the code chunks below is used to create an interactive point symbol map.

```{r}
tm_shape(mpsz_svy21)+
  tm_polygons() +
tm_shape(condo_resale.sf) +  
  tm_dots(col = "SELLING_PRICE",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

Notice that [`tm_dots()`](https://www.rdocumentation.org/packages/tmap/versions/2.2/topics/tm_symbols) is used instead of `tm_bubbles()`.`set.zoom.limits` argument of `tm_view()` sets the minimum and maximum zoom level to 11 and 14 respectively.

The code below will be used to turn R display into `plot` mode.

```{r}
tmap_mode("plot")
```

## Hedonic Pricing Modelling in R

### Simple Linear Regression Method

The code chunk below builds a simple linear regression model by using *SELLING_PRICE* as the dependent variable and *AREA_SQM* as the independent variable.

```{r}
condo.slr <- lm(formula=SELLING_PRICE ~ AREA_SQM, data = condo_resale.sf)
```

`lm()` returns an object of class "lm" or for multiple responses of class c("mlm", "lm").

The functions `summary()` and `anova()` can be used to obtain and print a summary and analysis of variance table of the results. The generic accessor functions coefficients, effects, fitted.values and residuals extract various useful features of the value returned by `lm`.

```{r}
summary(condo.slr)
```

The output report reveals that the SELLING_PRICE can be explained by using the formula:

          *y = -258121.1 + 14719x1*

The R-squared of 0.4518 reveals that the simple regression model built is able to explain about 45% of the resale prices. Since p-value is much smaller than 0.0001, I will reject the null hypothesis that mean is a good estimator of SELLING_PRICE. It could be inferred that simple linear regression model above is a good estimator of *SELLING_PRICE*.

The **Coefficients:** section of the report reveals that the p-values of both the estimates of the Intercept and ARA_SQM are smaller than 0.001. In view of this, the null hypothesis of the B0 and B1 are equal to 0 will be rejected. As a results, it could be inferred that the B0 and B1 are good parameter estimates.

To visualise the best fit curve on a scatterplot, the code chunk below incorporates `lm()` as a method function in ggplot's geometry.

```{r}
ggplot(data=condo_resale.sf,  
       aes(x=`AREA_SQM`, y=`SELLING_PRICE`)) +
  geom_point() +
  geom_smooth(method = lm)
```

Figure above reveals that there are a few statistical outliers with relatively high selling prices.

### Multiple Linear Regression Method

#### Visualising the relationships of the independent variables

Before building a multiple regression model, it is important to ensure that the indepdent variables used are not highly correlated to each other. If these highly correlated independent variables are used in building a regression model by mistake, the quality of the model will be compromised. This phenomenon is known as **multicollinearity** in statistics.

Correlation matrix is commonly used to visualise the relationships between the independent variables. Beside the `pairs()` of R, there are many packages support the display of a correlation matrix. In this section, the [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) package will be used.

The code chunk below is used to plot a scatterplot matrix of the relationship between the independent variables in *condo_resale* data.frame.

```{r, fig.width=10, fig.height=10}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")
```

Matrix reorder is very important for mining the hiden structure and patter in the matrix. There are four methods in corrplot (parameter order), named "AOE", "FPC", "hclust", "alphabet". In the code chunk above, AOE order is used. It orders the variables by using the *angular order of the eigenvectors* method suggested by [Michael Friendly](https://www.datavis.ca/papers/corrgram.pdf).

From the scatterplot matrix, it is clear that ***Freehold*** is highly correlated to ***LEASE_99YEAR***. In view of this, it is wiser to only include either one of them in the subsequent model building. As a result, ***LEASE_99YEAR*** is excluded in the subsequent model building.

### Building a hedonic pricing model using multiple linear regression method

The code chunk below using `lm()` to calibrate the multiple linear regression model.

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale.sf)
summary(condo.mlr)
```

### Preparing Publication Quality Table: olsrr method

With reference to the report above, it is clear that not all the independent variables are statistically significant. The model need to be recised by removing those variables which are not statistically significant.

The code chunk below calibrates the revised model.

```{r}
condo.mlr1 <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                 data=condo_resale.sf)
ols_regress(condo.mlr1)
```

### Preparing Publication Quality Table: gtsummary method

The [**gtsummary**](https://www.danieldsjoberg.com/gtsummary/) package provides an elegant and flexible way to create publication-ready summary tables in R.

In the code chunk below, [`tbl_regression()`](https://www.danieldsjoberg.com/gtsummary/reference/tbl_regression.html) is used to create a well formatted regression report.

```{r}
tbl_regression(condo.mlr1, intercept = TRUE)
```

With gtsummary package, model statistics can be included in the report by either appending them to the report table by using [`add_glance_table()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) or adding as a table source note by using [`add_glance_source_note()`](https://www.danieldsjoberg.com/gtsummary/reference/add_glance.html) as shown in the code chunk below.

```{r}
tbl_regression(condo.mlr1, 
               intercept = TRUE) %>% 
  add_glance_source_note(
    label = list(sigma ~ "\U03C3"),
    include = c(r.squared, adj.r.squared, 
                AIC, statistic,
                p.value, sigma))
```

#### Checking for multicolinearity

In this section, I would explore a R package specially programmed for performing OLS regression. It is called [**olsrr**](https://olsrr.rsquaredacademy.com/). It provides a collection of very useful methods for building better multiple linear regression models:

-   comprehensive regression output

-   residual diagnostics

-   measures of influence

-   heteroskedasticity tests

-   collinearity diagnostics

-   model fit assessment

-   variable contribution assessment

-   variable selection procedures

In the code chunk below, the [`ols_vif_tol()`](https://olsrr.rsquaredacademy.com/reference/ols_coll_diag.html) of **olsrr** package is used to test if there are sign of multicollinearity.

```{r}
ols_vif_tol(condo.mlr1)
```

Since the VIF of the independent variables are less than 10, it could be concluded that there are no sign of multicollinearity among the independent variables.

#### Test for Non-Linearity

In multiple linear regression, it is important to test the assumption that linearity and additivity of the relationship between dependent and independent variables.

In the code chunk below, the [`ols_plot_resid_fit()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_fit.html) of **olsrr** package is used to perform linearity assumption test.

```{r}
ols_plot_resid_fit(condo.mlr1)
```

The figure above reveals that most of the data poitns are scattered around the 0 line, hence it could be concluded that the relationships between the dependent variable and independent variables are linear.

#### Test for Normality Assumption

Lastly, the code chunk below uses [`ols_plot_resid_hist()`](https://olsrr.rsquaredacademy.com/reference/ols_plot_resid_hist.html) of *olsrr* package to perform normality assumption test.

```{r}
ols_plot_resid_hist(condo.mlr1)
```

The figure reveals that the residual of the multiple linear regression model (i.e. condo.mlr1) is resemble normal distribution.

If you prefer formal statistical test methods, the [`ols_test_normality()`](https://olsrr.rsquaredacademy.com/reference/ols_test_normality.html) of **olsrr** package can be used as shown in the code chunk below.

```{r}
ols_test_normality(condo.mlr1)
```

The summary table above reveals that the p-values of the four tests are way smaller than the alpha value of 0.05. Hence we will reject the null hypothesis that the residual is NOT resemble normal distribution.

#### Testing for Spatial Autocorrelation

The hedonic model we try to build are using geographically referenced attributes, hence it is also important for us to visual the residual of the hedonic pricing model.

In order to perform spatial autocorrelation test, we need to convert *condo_resale.sf* from sf data frame into a **SpatialPointsDataFrame**.

First, the code chunk below exports the residual of the hedonic pricing model and save it as a data frame.

```{r}
mlr.output <- as.data.frame(condo.mlr1$residuals)
```

Next, the code chunk below joins the newly created data frame with *condo_resale.sf* object.

```{r}
condo_resale.res.sf <- cbind(condo_resale.sf, 
                        condo.mlr1$residuals) %>%
rename(`MLR_RES` = `condo.mlr1.residuals`)
```

Next, the code chunk below converts *condo_resale.res.sf* from simple feature object into a SpatialPointsDataFrame because spdep package can only process sp conformed spatial data objects.

```{r}
condo_resale.sp <- as_Spatial(condo_resale.res.sf)
condo_resale.sp
```

Next, the code chunk below uses **tmap** package to display the distribution of the residuals on an interactive map.

```{r}
tmap_mode("view")
```

The code chunks below is used to create an interactive point symbol map.

```{r}
tm_shape(mpsz_svy21)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale.res.sf) +  
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style="quantile") +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

The figure above reveal that there is sign of spatial autocorrelation and hence the Moran's I test will be performed.

First, the code chunk below computes the distance-based weight matrix by using [`dnearneigh()`](https://r-spatial.github.io/spdep/reference/dnearneigh.html) function of **spdep**.

```{r}
nb <- dnearneigh(coordinates(condo_resale.sp), 0, 1500, longlat = FALSE)
summary(nb)
```

Next, [`nb2listw()`](https://r-spatial.github.io/spdep/reference/nb2listw.html) of **spdep** packge will be used to convert the output neighbours lists (i.e. nb) into a spatial weights.

```{r}
nb_lw <- nb2listw(nb, style = 'W')
summary(nb_lw)
```

Next, [`lm.morantest()`](https://r-spatial.github.io/spdep/reference/lm.morantest.html) of **spdep** package will be used to perform Moran's I test for residual spatial autocorrelation

```{r}
lm.morantest(condo.mlr1, nb_lw)
```

The Global Moran's I test for residual spatial autocorrelation shows that it's p-value is less than 0.00000000000000022 which is less than the alpha value of 0.05. Hence, I will reject the null hypothesis that the residuals are randomly distributed. Since the Observed Global Moran I = 0.1424418 which is greater than 0, it could be inferred that the residuals resemble cluster distribution.

## Building Hedonic Pricing Models using GWmodel

In this section, I will practice how to modelling hedonic pricing using both the fixed and adaptive bandwidth schemes.

### Building Fixed Bandwidth GWR Model

#### Computing fixed bandwith

In the code chunk below `bw.gwr()` of GWModel package is used to determine the optimal fixed bandwidth to use in the model. Notice that the argument ***adaptive*** is set to **FALSE** indicates that we are interested to compute the fixed bandwidth.

There are two possible approaches can be uused to determine the stopping rule, they are: CV cross-validation approach and AIC corrected (AICc) approach. We define the stopping rule using ***approach*** argeement.

```{r}
bw.fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                     PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                     PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                     PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                     FAMILY_FRIENDLY + FREEHOLD, 
                   data=condo_resale.sp, 
                   approach="CV", 
                   kernel="gaussian", 
                   adaptive=FALSE, 
                   longlat=FALSE)
```

The result shows that the recommended bandwidth is 971.3398 metres.

#### GWModel method - fixed bandwith

The code chunk below calibrates the gwr model using fixed bandwidth and gaussian kernel.

```{r}
gwr.fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + 
                         PROX_CHILDCARE + PROX_ELDERLYCARE  + PROX_URA_GROWTH_AREA + 
                         PROX_MRT   + PROX_PARK + PROX_PRIMARY_SCH + 
                         PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + 
                         FAMILY_FRIENDLY + FREEHOLD, 
                       data=condo_resale.sp, 
                       bw=bw.fixed, 
                       kernel = 'gaussian', 
                       longlat = FALSE)
```

The output is saved in a list of class "gwrm". The code below can be used to display the model output.

```{r}
gwr.fixed
```

The report shows that the adjusted r-square of the gwr is 0.8430 which is significantly better than the globel multiple linear regression model of 0.6472.

### Building Adaptive Bandwidth GWR Model

In this section, I will calibrate the gwr-absed hedonic pricing model by using adaptive bandwidth approach.

#### Computing the adaptive bandwidth

Similar to the earlier section, we will first .

The code chunk used use *bw.ger()* to determine the recommended data point to use by computing the adaptive bandwidth.

```{r}
bw.adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale.sp, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```

The result shows that the 30 is the recommended data points to be used.

#### Constructing the adaptive bandwidth gwr model

The code chunk below calibrates the gwr-based hedonic pricing model by using adaptive bandwidth and gaussian kernel.

```{r}
gwr.adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale.sp, bw=bw.adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE, 
                          longlat = FALSE)
```

The code below can be used to display the model output.

```{r}
gwr.adaptive
```

The report shows that the adjusted r-square of the gwr is 0.8561 which is significantly better than the globel multiple linear regression model of 0.6472.

### Visualising GWR Output

In addition to regression residuals, the output feature class table includes fields for observed and predicted y values, condition number (cond), Local R2, residuals, and explanatory variable coefficients and standard errors:

-   Condition Number: this diagnostic evaluates local collinearity. In the presence of strong local collinearity, results become unstable. Results associated with condition numbers larger than 30, may be unreliable.

-   Local R2: these values range between 0.0 and 1.0 and indicate how well the local regression model fits observed y values. Very low values indicate the local model is performing poorly. Mapping the Local R2 values to see where GWR predicts well and where it predicts poorly may provide clues about important variables that may be missing from the regression model.

-   Predicted: these are the estimated (or fitted) y values 3. computed by GWR.

-   Residuals: to obtain the residual values, the fitted y values are subtracted from the observed y values. Standardized residuals have a mean of zero and a standard deviation of 1. A cold-to-hot rendered map of standardized residuals can be produce by using these values.

-   Coefficient Standard Error: these values measure the reliability of each coefficient estimate. Confidence in those estimates are higher when standard errors are small in relation to the actual coefficient values. Large standard errors may indicate problems with local collinearity.

They are all stored in a SpatialPointsDataFrame or SpatialPolygonsDataFrame object integrated with fit.points, GWR coefficient estimates, y value, predicted values, coefficient standard errors and t-values in its "data" slot in an object called **SDF** of the output list.

### Converting SDF into *sf* data.frame

To visualise the fields in **SDF**, the code chunk below first coverts it into **sf** data.frame.

```{r}
condo_resale.sf.adaptive <- st_as_sf(gwr.adaptive$SDF) %>%
  st_transform(crs=3414)
condo_resale.sf.adaptive.svy21 <- st_transform(condo_resale.sf.adaptive, 3414)
condo_resale.sf.adaptive.svy21 
gwr.adaptive.output <- as.data.frame(gwr.adaptive$SDF)
condo_resale.sf.adaptive <- cbind(condo_resale.res.sf, as.matrix(gwr.adaptive.output))
```

Next, `glimpse()` is used to display the content of *condo_resale.sf.adaptive* sf data frame.

```{r}
glimpse(condo_resale.sf.adaptive)
```

```{r}
summary(gwr.adaptive$SDF$yhat)
```

### Visualising local R2

The code chunks below is used to create an interactive point symbol map.

```{r}
tmap_mode("view")
tm_shape(mpsz_svy21)+
  tm_polygons(alpha = 0.1) +
tm_shape(condo_resale.sf.adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_mode("plot")
```

#### By URA Planning Region

```{r}
tm_shape(mpsz_svy21[mpsz_svy21$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(condo_resale.sf.adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```
